Cherenkov radiation and scattering of external dispersive waves by two-color solitons

### 2022-07-07 text/Draft.tex

\documentclass[aps, prl, twocolumn, superscriptaddress, final]{revtex4}
\pdfoutput=1

@@

\author{Oliver Melchert}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Institute of Quantum Optics, Leibniz Universit\"at Hannover, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Hannover Centre for Optical Technologies, Neinburger Strasse 17, 30167 Hannover, Germany}

\author{Stephanie Willms}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Institute of Quantum Optics, Leibniz Universit\"at Hannover, Welfengarten 1, 30167 Hannover, Germany}

\author{Surajit Bose}
\affiliation{Institute of Quantum Optics, Leibniz Universit\"at Hannover, Welfengarten 1, 30167 Hannover, Germany}

\author{Ihar Babushkin}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Institute of Quantum Optics, Leibniz Universit\"at Hannover, Welfengarten 1, 30167 Hannover, Germany}

\author{Ayhan Demircan}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Institute of Quantum Optics, Leibniz Universit\"at Hannover, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Hannover Centre for Optical Technologies, Neinburger Strasse 17, 30167 Hannover, Germany}

\author{Alexey Yulin}
\affiliation{Department of Nanophotonics and Metamaterials, ITMO University, Kronverskiy pr. 49, 19701 St.~Petersburg, Russia}

@@

\begin{abstract}
  In dispersion landscapes with two separate regions of anomalous dispersion it is possible to create a quasi-stable two-color solitary wave. In this paper we consider how those waves interact with dispersive radiation, including the processes of Cherenkov radiation generation as well as the processes of external wave scattering. We derive the analytic resonance conditions and verify them through numeric experiments.
\end{abstract}

@@

From the prior work~\cite{melchert2019soliton} we know that if the dispersion landscape $\beta(\omega)$ has two regions of anomalous dispersion separated by a region of normal dispersion, it is possible to create a quasi-stable configuration of two tightly-coupled solitons each propagating on its own carrier frequency. This coupled state, especially in the process of initial evolution, sheds dispersive waves that resembles Cherenkov radiation that is observed for other types of solitary waves in vast variety of settings~\cite{akhmediev1995cherenkov, afanasjev1996effect, driben2015resonant, conforti2015parametric, wright2015ultrabroadband, oreshnikov2017dispersive}.

{
  % This is just a dummy text to see whether the general layout of the paper makes sense.
  \it
Nullam eu ante vel est convallis dignissim.  Fusce suscipit, wisi nec facilisis facilisis, est dui fermentum leo, quis tempor ligula erat quis odio.  Nunc porta vulputate tellus.  Nunc rutrum turpis sed pede.  Sed bibendum.  Aliquam posuere.  Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis varius mi purus non odio.  Pellentesque condimentum, magna ut suscipit hendrerit, ipsum augue ornare nulla, non luctus diam neque sit amet urna.  Curabitur vulputate vestibulum lorem.  Fusce sagittis, libero non molestie mollis, magna orci ultrices dolor, at vulputate neque nulla lacinia eros.  Sed id ligula quis est convallis tempor.  Curabitur lacinia pulvinar nibh.  Nam a sapien.

Pellentesque dapibus suscipit ligula.  Donec posuere augue in quam.  Etiam vel tortor sodales tellus ultricies commodo.  Suspendisse potenti.  Aenean in sem ac leo mollis blandit.  Donec neque quam, dignissim in, mollis nec, sagittis eu, wisi.  Phasellus lacus.  Etiam laoreet quam sed arcu.  Phasellus at dui in ligula mollis ultricies.  Integer placerat tristique nisl.  Praesent augue.  Fusce commodo.  Vestibulum convallis, lorem a tempus semper, dui dui euismod elit, vitae placerat urna tortor vitae lacus.  Nullam libero mauris, consequat quis, varius et, dictum id, arcu.  Mauris mollis tincidunt felis.  Aliquam feugiat tellus ut neque.  Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia purus, et dictum nunc justo sit amet elit.
}

\section{Analytic model}%
\label{sec:AnalyticModel}

In this section we derive a perturbation theory that explains the resonance conditions that were proposed before, and then we extend to it to describe the process of external waves scattering on the soliton. We consider a non-envelope version of a nonlinear Schr\"odinger equation~\cite{amiranashvili2010hamiltonian}
\begin{equation}
  \label{eq:GNLSE}
  i \partial_{z} \tilde u
    + \beta(\omega) \tilde u(z, \omega)
    + \gamma(\omega) \mathcal{F} \left\{
      \abs{u}^{2} u(z, t)
    \right\}_{(\omega > 0)} = 0.
\end{equation}
Here and further tilde indicates a Fourier image of the field and $\mathcal{F}\left\{ \cdots \right\}_{(\omega > 0)}$ is the explicit Fourier transform over the positive frequencies.

@@

\noindent We start by introducing the ansatz that represents the solution as a sum of two single-frequency solitary waves $U_{1, 2}(z, t)$ and a residue radiation $\psi(z, t)$
\begin{gather}
  \label{eq:PerturbationAnsatz}
  u(z, t) = U_{1}(z, t) + U_{2}(z, t) + \psi(z, t) \\
  \abs{\psi} \ll \abs{U_{1}} \sim \abs{U_{2}}. \nonumber
\end{gather}
For the solitary waves $U_{1, 2}(z, t)$ we assume that they satisfy a pair of coupled nonlinear Schr\"odinger equations below
\begin{equation}
  \label{eq:CoupledSolitons}
  i \partial_{z} U_{n}
    + \beta_{n}(i \partial_{t}) U_{n}(z, t)
    + \gamma(\omega_{n}) \left(
      \abs{U_{n}}^{2} U_{n} + 2 \abs{U_{m}}^{2} U_{n}
    \right) = 0,
\end{equation}
where $n = 1, 2$ and $m = 2, 1 \ne n$. Essentially, this is an assumption that the solitons are coupled strictly by inducing a refractive-index potential on each other, and each soliton exists in some sort of a truncated dispersion landscape that is defined by the operator $\beta_{n}(i \partial_{t})$. A reasonable guess for a truncated operator is a parabolic approximation close to the carrier frequency
\begin{equation}
  \label{eq:TruncatedDispersionOperator}
  \beta_{n}(i \partial_{t}) =
    \beta(\omega_{n}) +
    i \beta'(\omega_{n}) \partial_{t} -
    \frac{1}{2} \, \beta''(\omega_{n}) \partial_{t}^{2}
\end{equation}
We additionally propose that in the frame of reference of the individual soliton the envelope evolves with wavenumber $k_{n}(\omega)$ which we approximate by the wavenumber of the fundamental soliton in the ordinary nonlinear Schr\"odinger equation and a correction from a secondary soliton as
\begin{equation}
  \label{eq:SolitonWavenumber}
  k_{n}(\omega)
    \approx \frac{\gamma(\omega_{n}) A_{n}^{2}}{2}
    + \beta(\omega_{n}) + \beta'(\omega_{n}) (\omega - \omega_{n})
    + \gamma(\omega_{n}) A_{m}^{2}
\end{equation}
where $n = 1, 2$, $m = 2, 1 \ne n$ and $A_{n}$ is the soliton amplitude.

By substituting ansatz~\eqref{eq:PerturbationAnsatz} into equation~\eqref{eq:GNLSE}, linearizing with respect to perturbation $\psi(z, \omega)$ and discarding the terms corresponding to the soliton equations~\eqref{eq:CoupledSolitons} we get the equation for $\tilde \psi$
\begin{multline}
  \label{eq:PerturbationEquation}
  i \partial_{z} \tilde \psi
    + \beta(\omega) \tilde \psi(z, \omega)
    + \gamma(\omega) \mathcal{F}\left\{
      2 \abs{U_{1} + U_{2}}^{2} \psi +
      \left( U_{1} + U_{2} \right)^{2} \psi^{*}
    \right\}_{(\omega > 0)} = \\
    - \left[ \beta(\omega) - \beta_{1}(\omega) \right] \tilde U_{1}(z, \omega)
    - \left[ \beta(\omega) - \beta_{2}(\omega) \right] \tilde U_{2}(z, \omega)
    - \gamma(\omega) \mathcal{F} \left\{
      U_{2}^{2} U_{1}^{*} + U_{1}^{2} U_{2}^{*}
    \right\}_{(\omega > 0)}
\end{multline}
In the right-hand's side of equation~\eqref{eq:PerturbationEquation} we see two types of driving terms and each of those terms can be in resonance with the linear dispersive waves that exist in the system, if the particular wavenumber $k(\omega_{*})$ of the driving term is equal to a wavenumber of a dispersive wave $\beta(\omega_{*})$ at some frequency $\omega_{*}$~\cite{akhmediev1995cherenkov, yulin2004four}. The first type of the terms is $\left[ \beta(\omega) - \beta_{n}(\omega) \right] U_{n}(z, \omega)$ which drives the generation of Cherenkov radiation by an individual soliton $U_{n}$ if the following resonance condition is satisfied
\begin{equation}
  \label{eq:CherenkovRadiationResonanceCondition}
  \beta(\omega) = k_{n}(\omega).
\end{equation}
The second type of the terms are $\gamma(\omega) \mathcal{F}\left\{ U_{2}^{2} U_{1}^{*} \right\}$ and $\gamma(\omega) \mathcal{F}\left\{ U_{1}^{2} U_{2}^{*} \right\}$ and they correspond to the process of four-wave mixing that in our case results in generation of dispersive radiation at some frequency where
\begin{equation}
  \label{eq:FWMRadiationResonanceCondition}
  \beta(\omega) = 2 k_{n}(\omega)  - k_{m}(\omega).
\end{equation}

Let us move on to the problem of external dispersive wave scattering. For that we split the perturbation into the incident and the scattered parts
\begin{equation}
  \label{eq:PerturbationSplit}
  \psi(z, t) = \psi_{\text{inc}}(z, t) + \psi_{\text{sc}}(z, t).
\end{equation}
We explicitly define $\psi_{\text{inc}}(z, t)$ as a linear wave that is propagating in a soliton-free medium
\begin{equation}
  \label{eq:IncidentFieldEquation}
  i \partial_{z} \tilde \psi_{\text{inc}}(z, t)
    + \beta(\omega) \tilde \psi_{\text{inc}}(z, t) = 0.
\end{equation}
Substituting equation~\eqref{eq:PerturbationSplit} into~\eqref{eq:PerturbationEquation} and eliminating the terms corresponding to~\eqref{eq:IncidentFieldEquation} we are left with equation for the scattered component
\begin{multline}
  \label{eq:ScatteredFieldEquation}
  i \partial_{z} \tilde \psi_{\text{sc}}
    + \beta(\omega) \tilde \psi_{\text{sc}}(z, \omega)
    + \gamma(\omega) \mathcal{F}\left\{
      2 \abs{U_{1} + U_{2}}^{2} \psi_{\text{sc}} +
      \left( U_{1} + U_{2} \right)^{2} \psi_{\text{sc}}^{*}
    \right\}_{(\omega > 0)} = \text{\ldots (omitted is the RHS from~\eqref{eq:PerturbationEquation})} \\
    - \gamma(\omega) \mathcal{F}\left\{
      2 \left(
        \abs{U_{1}}^{2} + U_{1} U_{2}^{*} + U_{2} U_{1}^{*} + \abs{U_{2}}^{2}
      \right) \psi_{\text{inc}} +
      \left(
        U_{1}^{2} + 2 \, U_{1} U_{2} + U_{2}^{2}
      \right) \psi_{\text{inc}}^{*}
    \right\}_{(\omega > 0)}
\end{multline}
\end{widetext}

In addition to the resonance terms already discussed in equation~\eqref{eq:PerturbationEquation} we see the terms that arise due to interaction between the incident radiation and the soliton. In here six new types of resonance behaviour are possible. The first one is due to terms $\abs{U_{1}}^{2} \psi_{\text{inc}}$ and $\abs{U_{2}}^{2} \psi_{\text{inc}}$, both with the resonance condition
\begin{equation}
  \label{eq:Type0ResonanceCondition}
  \beta(\omega_{\text{sc}}) = \beta(\omega_{\text{inc}})
\end{equation}
The next two are due to mixed terms $U_{1} U_{2}^{*} \psi_{\text{inc}}$ and $U_{2} U_{1}^{*} \psi_{\text{inc}}$, with the corresponding resonance condition being
\begin{equation}
  \label{eq:Type1ResonanceCondition}
  \beta(\omega_{\text{sc}}) = \pm k_{1} \mp k_{2} + \beta(\omega_{\text{inc}})
\end{equation}
Another two are due to $U_{n}^{2} \psi_{\text{inc}}^{*}$ and the resonance condition is
\begin{equation}
  \label{eq:Type2ResonanceCondition}
  \beta(\omega_{\text{sc}}) =
    2 k_{n}(\omega_{\text{sc}}) -
    \beta(\omega_{\text{inc}}), ~ n = 1, 2.
\end{equation}
The final one is due to $2 U_{1} U_{2} \psi_{\text{inc}}^{*}$ with the resonances at
\begin{equation}
  \label{eq:Type3ResonanceCondition}
  \beta(\omega_{\text{sc}}) =
    k_{1}(\omega_{\text{sc}}) +
    k_{2}(\omega_{\text{sc}}) -
    \beta(\omega_{\text{inc}}).
\end{equation}
To verify the predictions given by resonance conditions~\eqref{eq:CherenkovRadiationResonanceCondition},~\eqref{eq:FWMRadiationResonanceCondition},~\eqref{eq:Type0ResonanceCondition},~\eqref{eq:Type1ResonanceCondition},~\eqref{eq:Type2ResonanceCondition}, and~\eqref{eq:Type3ResonanceCondition} we proceed to numeric experiments.

\section{Numeric experiments}%
\label{sec:NumericExperiments}

To numerically integrate equation~\eqref{eq:GNLSE} we use the integrating factor method and transform the equation into a non-stiff version for a modified spectrum~\cite{dudley2010supercontinuum}. The modified equation can be handled by any standard ODE solver; we use a \texttt{scipy} interface to \texttt{ZVODE} solver from \texttt{ODEPACK}~\cite{hindmarsh1983odepack, virtanen2020scipy} (the complete set of sources codes used in the paper can be found in~\cite{sources}).

>>>>

\documentclass[aps, pra, twocolumn, superscriptaddress, final]{revtex4}
\pdfoutput=1

@@

\author{Oliver Melchert}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Quantum Optics, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Hannover Centre for Optical Technologies, Nienburger Strasse 17, 30167 Hannover, Germany}

\author{Stephanie Willms}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Quantum Optics, Welfengarten 1, 30167 Hannover, Germany}

\author{Surajit Bose}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Photonics, Nienburger Strasse 17, 30167 Hannover, Germany}

\author{Ihar Babushkin}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Quantum Optics, Welfengarten 1, 30167 Hannover, Germany}

\author{Ayhan Demircan}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Quantum Optics, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Hannover Centre for Optical Technologies, Nienburger Strasse 17, 30167 Hannover, Germany}

\author{Uwe Morgner}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Quantum Optics, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Hannover Centre for Optical Technologies, Nienburger Strasse 17, 30167 Hannover, Germany}

\author{Alexey Yulin}
\affiliation{Department of Physics, ITMO University, Kronverskiy pr. 49, 19701 St.~Petersburg, Russia}

@@

\begin{abstract}
For waveguides with two separate regions of anomalous dispersion, it is
possible to create a quasi-stable two-color solitary wave.
In this paper we consider how those waves interact with dispersive radiation,
both generation of Cherenkov radiation and scattering of incident dispersive
waves.
We derive the analytic resonance conditions and verify them through numeric
experiments. We also report incident radiation driving the internal
oscillations of the soliton during the scattering process in case of an intense
incident radiation.
We generalize the resonance conditions for the case of an oscillating soliton
and demonstrate how one can use the scattering process to probe and excite an
internal mode of two-color soliton molecules.
\end{abstract}

@@

Solitary waves or solitons are localized nonlinear excitations that preserve
their shapes during evolution. In the systems close to integrable ones solitons
are proven to be robust and in many cases different interactions result
only in the variation of the soliton parameters, such as its intensity or
frequency.
In a certain sense solitons can be called ``eigenmodes of the nonlinear
problem'', meaning that the dynamics of the system can be considered as a set
of solitons interacting with quasi-linear dispersive waves.
Due to their robustness, solitons are of fundamental as well as practical
importance, for example, in the context of optical supercontinuum
generation~\cite{RevModPhys.78.1135, RevModPhys.82.1287,
dudley2010supercontinuum} or soliton fiber
lasers~\cite{Mollenauer:84,Kafka:89,Matsas_1992, Turitsyn_2016}.

The mutual interaction of optical solitons in fibers can enable the formation
of bound states, often referred to as soliton molecules.
%
They can be realized via dispersion engineering in the framework of the
standard nonlinear Schrödinger equation (NSE) and consist of two pulses that
maintain a fixed separation in time \cite{Stratmann:PRL:2005,Hause:PRA:2008}.
%
Further, optical soliton clusters have been discovered and studied in a large
variety of physical systems described by different equations such as the
generalized nonlinear Schr\"odinger equation~\cite{PhysRevE.51.3572,
PhysRevE.47.2874, PhysRevA.80.043829}, coupled NLSs~\cite{Kivshar:89,
Abdullaev:89}, the Ginzburg-Landau
equation~\cite{PhysRevLett.79.4047, Haelterman:97}, Lugiato-Lefever
equation~\cite{Weng_LugiatoLefever} and many others~\cite{Neshev:03,
Afanasjev:94, PhysRevE.67.056610, Chen:00, PhysRevLett.89.273902,
PhysRevLett.122.084101, PhysRevLett.121.023905, PhysRevLett.118.243901,
PhysRevE.86.036606, Grelu_NatPhoton, TURMANOV20151828}.  The aforementioned
bound states of solitons have a single central frequency and the whole spectrum
is localized around this frequency.

Another possibility to observe soliton molecules is to provide interaction
between solitons having their carrier frequencies well detuned from each
other. To make the interaction efficient the velocities of the solitons must be
close. In the case of scalar solitons this means that we need higher order
dispersion with two different frequency ranges where the solitons can be
accommodated. These solitons were recently reported
in~\cite{PhysRevLett.123.243905, PhysRevA.101.043822, Melchert:21,
Melchert_SciRep21} for pulses propagating in conservative fibers with higher
order dispersion. Similar solitons were discovered also in mode-locked cavity
lasers~\cite{Lourdesamy_NaturePhys} and in coherently pumped ring
resonators~\cite{Moille:18, Melchert:20}.

In the case of nonintegrable systems, solitons can interact with dispersive
waves (DWs) of low intensity and this interaction leads to interesting physical
phenomena, as the efficient generation of new frequencies~\cite{Yulin:04,
PhysRevE.72.016619, Efimov:04, PhysRevLett.95.213902, Efimov:06}.
It should be noted that this effect is closely related to the optical push
broom effect~\cite{deSterke:92}.
Such resonant scattering can be cascaded, enriching optical
supercontinuum spectra significantly~\cite{Gorbach_NatPhoton}.
Similar effects
were studied in Refs.~\cite{PhysRevLett.110.233901,  Demircan_SciRep,
Philbin_Science}, where the term ``optical event horizon'' was coined. It was
also established that resonant dispersive waves affect the parameters of the
solitons and can lead to dispersive wave mediated acceleration of
solitons~\cite{PhysRevE.72.016619, PhysRevLett.106.163901,
PhysRevLett.110.233901, Yulin:13, Demircan:14,  Wang:15, PhysRevA.98.023833}.
It should be mentioned that resonant scattering of dispersive waves is also
studied for dark solitons~\cite{Oreshnikov:15, Marest:18,  Deng:18,
PhysRevA.103.023505}, and oscillating solitons~\cite{Kodama:94,
Conforti_SciRep, Driben:15, PhysRevLett.115.223902, PhysRevLett.116.183901,
PhysRevA.96.013809}.

The present paper aims to study in detail the interaction of dispersive waves
with two-color soliton molecules, consisting of two bound solitons having well separated
frequencies.
%
In the prior work~\cite{PhysRevLett.123.243905} we have demonstrated how to
create quasi-stable configurations of two tightly-coupled pulses in a
dispersion landscape $\beta(\omega)$ with two regions of anomalous dispersion,
separated by a region of normal dispersion. Each of these subpulses propagates
on its own carrier frequency.
%
This coupled state, especially in the process of initial
evolution, sheds dispersive waves that resemble Cherenkov radiation that is
observed for other types of solitary waves in a vast variety of
settings~\cite{akhmediev1995cherenkov, afanasjev1996effect, Driben:15,
Conforti_SciRep, PhysRevLett.115.223902, PhysRevA.96.013809}. And since DW
generation is demonstrated, it is reasonable to anticipate that two-color
soliton molecules colliding with DWs will produce resonant emission in the same
way as it occurs in the case of conventional solitons.
Thanks to a more complex structure of the two-color solitons, one can expect
the scattering dynamics to be much richer compared to the conventional single
soliton case. For instance, it is possible that the internal degrees of freedom
can be excited due to the interaction with a DW and this should
strongly affect the dynamics of the system.
Subsequently we confirm this influence, providing a combined analytical and
numerical view of the process.

The paper is structured as follows.
In Sect.~\ref{sec:analytic_model} we introduce a mathematical model for the
light propagating in a nonlinear fiber with higher order dispersion and derive
the condition of resonant four-wave mixing of the soliton molecules with DWs.
In Sect.~\ref{sec:num_exp} we report the results of the numerical simulations
of the different propagation regimes of bichromatic soliton molecules. The
results of the simulations are compared against analytical resonance
conditions.
Section~\ref{sec:nonlin_eff} discusses the weakly nonlinear case where the
intensity of the dispersive waves is sufficiently large to modify the
interaction between the soliton molecules and the waves.  The excitation of the
internal mode is also discussed in this section.
The paper concludes with a summary in Sect.~\ref{sec:conclusions}.

\section{Analytical model\label{sec:analytic_model}}%

In this section we derive a perturbation theory that explains the resonance
conditions that were proposed before in \cite{PhysRevLett.123.243905},
demonstrate additional Cherenkov radiation mechanisms due to four-wave mixing
(FWM) between the frequency components of the soliton, and then extend this
theory to describe the process of scattering of external waves on the soliton.

We start by considering a non-envelope version of a nonlinear Schr\"odinger
equation~\cite{amiranashvili2010hamiltonian}
\begin{equation}
  \label{eq:GNLSE}
  i \partial_{z} \tilde u
    + \beta(\omega) \tilde u
    + \gamma(\omega) \mathcal{F} \left\{
      \abs{u}^{2} u
    \right\}_{(\omega > 0)} = 0.
\end{equation}
Here and further, $\tilde{u}\equiv \tilde{u}(z,\omega)$ indicates the Fourier
image of the field $u\equiv u(z,t)$ and $\mathcal{F}\left\{ \cdots
\right\}_{(\omega > 0)}$ is the explicit Fourier transform taken for the positive
frequencies only.

@@

\noindent To build the perturbation theory let us introduce an ansatz that
represents the solution as a sum of two single-frequency solitary waves $U_{1,
2}(z, t)$ and a small residue radiation $\psi(z, t)$
\begin{gather}
  \label{eq:PerturbationAnsatz}
  u(z, t) = U_{1}(z, t) + U_{2}(z, t) + \psi(z, t), \\
  \abs{\psi} \ll \abs{U_{1}} \sim \abs{U_{2}}. \nonumber
\end{gather}
For the solitary waves $U_{1, 2}(z, t)$ we assume that they satisfy a pair of
coupled nonlinear Schr\"odinger equations below
\begin{equation}
  \label{eq:CoupledSolitons}
  i \partial_{z} U_{n}
    + \beta_{n}(i \partial_{t}) U_{n}(z, t)
    + \gamma(\omega_{n}) \left(
      \abs{U_{n}}^{2} U_{n} + 2 \abs{U_{m}}^{2} U_{n}
    \right) = 0,
\end{equation}
where $n = 1, 2$ and $m = 2, 1 \ne n$. Essentially, this is an assumption that
the two-color soliton molecule consists of two pulses, which are incoherently
coupled by inducing a refractive-index potential on each
other~\cite{agrawal2013nonlinear}.  Each subpulse exists in some
sort of truncated dispersion landscape, defined by the operator $\beta_{n}(i
\partial_{t})$.
Under specific conditions, the subpulses are given by fundamental solitons of a
modified NLS \cite{Melchert:21}.
A reasonable guess for the truncated operator is a parabolic approximation close
to the carrier frequency
\begin{equation}
  \label{eq:TruncatedDispersionOperator}
  \beta_{n}(i \partial_{t}) =
    \beta(\omega_{n}) +
    i \beta'(\omega_{n}) \partial_{t} -
    \frac{1}{2} \, \beta''(\omega_{n}) \partial_{t}^{2}.
\end{equation}
We additionally suppose that in the soliton's frame of reference the envelope evolves with wavenumber $k_{n}(\omega)$ which we approximate by the wavenumber of the fundamental soliton in the ordinary nonlinear Schr\"odinger equation and a correction from a secondary soliton as
\begin{equation}
  \label{eq:SolitonWavenumber}
  k_{n}(\omega)
    \approx \frac{\gamma(\omega_{n}) A_{n}^{2}}{2}
    + \beta(\omega_{n}) + \beta'(\omega_{n}) (\omega - \omega_{n})
    + \gamma(\omega_{n}) A_{m}^{2},
\end{equation}
where $n = 1, 2$, $m = 2, 1 \ne n$ and $A_{n}$ is the soliton amplitude.

By substituting ansatz~\eqref{eq:PerturbationAnsatz} into
Eq.~\eqref{eq:GNLSE}, linearizing with respect to perturbation $\psi(z,
\omega)$, and discarding the terms corresponding to the soliton
Eqs.~\eqref{eq:CoupledSolitons}, we get the equation for $\tilde \psi$
\begin{multline}
  \label{eq:PerturbationEquation}
  i \partial_{z} \tilde \psi
    + \beta(\omega) \tilde \psi
    + \gamma(\omega) \mathcal{F}\left\{
      2 \abs{U_{1} + U_{2}}^{2} \psi +
      \left( U_{1} + U_{2} \right)^{2} \psi^{*}
    \right\}_{(\omega > 0)} = \\
    - \left[ \beta(\omega) - \beta_{1}(\omega) \right] \tilde U_{1}
    - \left[ \beta(\omega) - \beta_{2}(\omega) \right] \tilde U_{2}
    - \gamma(\omega) \mathcal{F} \left\{
      U_{2}^{2} U_{1}^{*} + U_{1}^{2} U_{2}^{*}
    \right\}_{(\omega > 0)}.
\end{multline}
On the right-hand side (RHS) of Eq.~\eqref{eq:PerturbationEquation} we see two
types of driving terms and each of those terms can be in resonance with the
linear DWs that exist in the system if the particular wavenumber
$k(\omega_{*})$ of the driving term is equal to the wavenumber of a DW
$\beta(\omega_{*})$ at some frequency
$\omega_{*}$~\cite{akhmediev1995cherenkov, Yulin:04}. The first type
is given by $\left[ \beta(\omega) - \beta_{n}(\omega) \right] \tilde U_{n}(z,
\omega)$ which drives the generation of Cherenkov radiation by an individual
soliton $U_{n}$ if the following resonance condition is satisfied
\begin{equation}
  \label{eq:CherenkovRadiationResonanceCondition}
  \beta(\omega) = k_{n}(\omega).
\end{equation}
The second type of the terms are $\gamma(\omega) \mathcal{F}\left\{ U_{2}^{2}
U_{1}^{*} \right\}$ and $\gamma(\omega) \mathcal{F}\left\{ U_{1}^{2} U_{2}^{*}
\right\}$ and they correspond to the process of four-wave mixing that in our
case results in the generation of dispersive radiation at some frequency where
\begin{equation}
  \label{eq:FWMRadiationResonanceCondition}
  \beta(\omega) = 2 k_{n}(\omega)  - k_{m}(\omega).
\end{equation}

Let us move on to the problem of external dispersive wave scattering. For that,
we split the perturbation into the incident and the scattered parts
\begin{equation}
  \label{eq:PerturbationSplit}
  \psi(z, t) = \psi_{\text{inc}}(z, t) + \psi_{\text{sc}}(z, t).
\end{equation}
We explicitly define $\psi_{\text{inc}}(z, t)$ as a linear wave that is propagating in a soliton-free medium
\begin{equation}
  \label{eq:IncidentFieldEquation}
  i \partial_{z} \tilde \psi_{\text{inc}}(z, t)
    + \beta(\omega) \tilde \psi_{\text{inc}}(z, t) = 0.
\end{equation}
Substituting Eq.~\eqref{eq:PerturbationSplit}
into~\eqref{eq:PerturbationEquation} and eliminating the terms corresponding to
Eq.~\eqref{eq:IncidentFieldEquation} we are left with the equation for the
scattered component
\begin{multline}
  \label{eq:ScatteredFieldEquation}
  i \partial_{z} \tilde \psi_{\text{sc}}
    + \beta(\omega) \tilde \psi_{\text{sc}}
    + \gamma(\omega) \mathcal{F}\left\{
      2 \abs{U_{1} + U_{2}}^{2} \psi_{\text{sc}} +
      \left( U_{1} + U_{2} \right)^{2} \psi_{\text{sc}}^{*}
    \right\}_{(\omega > 0)} = \text{\ldots (omitted is the RHS of Eq.~\eqref{eq:PerturbationEquation})} \\
    - \gamma(\omega) \mathcal{F}\left\{
      2 \left(
        \abs{U_{1}}^{2} + U_{1} U_{2}^{*} + U_{2} U_{1}^{*} + \abs{U_{2}}^{2}
      \right) \psi_{\text{inc}} +
      \left(
        U_{1}^{2} + 2 \, U_{1} U_{2} + U_{2}^{2}
      \right) \psi_{\text{inc}}^{*}
    \right\}_{(\omega > 0)}.
\end{multline}
\end{widetext}

In addition to the resonance terms already discussed in
Eq.~\eqref{eq:PerturbationEquation}, we see terms that arise due to
interaction between the incident radiation and the soliton. Here, six new
types of resonance behavior are possible. The first one is due to terms
$\abs{U_{1}}^{2} \psi_{\text{inc}}$ and $\abs{U_{2}}^{2} \psi_{\text{inc}}$,
both with the resonance condition
\begin{equation}
  \label{eq:Type0ResonanceCondition}
  \beta(\omega_{\text{sc}}) = \beta(\omega_{\text{inc}}).
\end{equation}
The next two are due to mixed terms $U_{1} U_{2}^{*} \psi_{\text{inc}}$ and $U_{2} U_{1}^{*} \psi_{\text{inc}}$, with the corresponding resonance condition being
\begin{equation}
  \label{eq:Type1ResonanceCondition}
  \beta(\omega_{\text{sc}}) = \pm k_{1} \mp k_{2} + \beta(\omega_{\text{inc}}).
\end{equation}
Another two are due to $U_{n}^{2} \psi_{\text{inc}}^{*}$ and the resonance condition is
\begin{equation}
  \label{eq:Type2ResonanceCondition}
  \beta(\omega_{\text{sc}}) =
    2 k_{n}(\omega_{\text{sc}}) -
    \beta(\omega_{\text{inc}}), ~ n = 1, 2.
\end{equation}
The final one is due to $2 U_{1} U_{2} \psi_{\text{inc}}^{*}$ with the resonances at
\begin{equation}
  \label{eq:Type3ResonanceCondition}
  \beta(\omega_{\text{sc}}) =
    k_{1}(\omega_{\text{sc}}) +
    k_{2}(\omega_{\text{sc}}) -
    \beta(\omega_{\text{inc}}).
\end{equation}
To verify the predictions given by resonance
conditions~\eqref{eq:CherenkovRadiationResonanceCondition},~\eqref{eq:FWMRadiationResonanceCondition},~\eqref{eq:Type0ResonanceCondition},~\eqref{eq:Type1ResonanceCondition},~\eqref{eq:Type2ResonanceCondition},
and~\eqref{eq:Type3ResonanceCondition} we proceed to numeric experiments.

\section{Numerical experiments\label{sec:num_exp}}%

To numerically integrate Eq.~\eqref{eq:GNLSE} we use the integrating factor
method and transform the equation into a non-stiff version for a modified
spectrum~\cite{dudley2010supercontinuum}. The modified equation can be handled
by any standard ODE solver; we use a \texttt{scipy} interface to \texttt{ZVODE}
solver from \texttt{ODEPACK}~\cite{hindmarsh1983odepack, virtanen2020scipy}
(the code necessary to reproduce the results in the paper can be found
in~\cite{sources}). All the computations are performed in a frame of reference
co-moving with the soliton, which is achieved by a transformation
\begin{equation*}
  t \to t - \beta'(\omega_{1}) z,
\end{equation*}
which, in turn, results in
\begin{align*}
  \beta(\omega) &\to
    \beta(\omega)
    - \beta(\omega_{1})
    - \beta'(\omega_{1}) (\omega - \omega_{1}) \\
  k_{1}(\omega) &\to
    \frac{\gamma(\omega_{1}) A_{1}^{2}}{2} + \gamma(\omega_{1}) A_{2}^{2} \\
  k_{2}(\omega) &\to
    \frac{\gamma(\omega_{2}) A_{2}^{2}}{2}
    + \gamma(\omega_{1}) A_{1}^{2}
    + \beta(\omega_{2}) - \beta(\omega_{1}).
\end{align*}
Description of the specific dispersion profile model $\beta(\omega)$ used in the
simulations can be found in the Supplemental Material~\cite{suppMat}.
